Secure Composition of Robust and Optimising Compilers

### 2023-07-09 popl/main.tex
parts of sec 5

\subsection{Shared Language Definitions}\label{subsec:cs:defs}
\begin{gather*}
  \begin{aligned}
  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}
  &
  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
  &
  \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
  \end{aligned}
\end{gather*}
\MPin{
	this latex had many issues.
	the \& were at different spots
	it used hspace to space things instead of \&
	and it did not use \& to align the following things on a line
	and it did not use multiple aligned into the gather
	(what's the point of gather if you don't have several aligned inside? use a single align* instead)
}

Above is a fraction of the shared syntax of all the programming languages of this paper.
\MP{why a fraction? what exprs are missing}
The syntax and semantics are mostly standard.\MP{and therefore omitted}
However, when allocating pointers\MP{arrays?} with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects\MP{cells?} and $\expr[_{2}]$ the overall size of an array.

\MPin{
	below is semantics info. should go with their symbols and the semantics rules.
	and a better description of the semantics symbols.

}
The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
Because of the sandboxing, locations pointing into heaps must be associated with a control--tag\MP{symbol} to know the heap they point into.
Other metadata in $\Omega$ entails a list of callable objects\MP{functions? objects are other things in PL} with information on wether they belong to the context or the component.
That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
For the languages considered in this paper, no callable objects are first-class.
So, the only values are pairs, pointers, and natural numbers.
Another non-standard part of the languages is the concept of the poison-tag.
The poison-tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$.
\MPin{
	last line is mixing operational semantics and trace semantics
}

@@

Possible events that are emitted during execution are similar to the ones of the specification trace model \MP{ref?}.
The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.\MP{why? to help with proofs?}
Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
The environmental semantics emits the tag $\nocomm$ for calls from context- to context-level or component- to component-level.
In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.
\MP{intuition? these are aux functions for the semantics}

>>>>

\subsection{Shared Language Definitions}\label{subsec:cs:defs}
\begin{gather*}
  \begin{aligned}
  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Functions)}~&\function\bnfdef\ \lfunction{g}{x}{e}
  &
  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
  &
  \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
  \end{aligned}
\end{gather*}
% \MPin{
%   this latex had many issues.
%   the \& were at different spots
%   it used hspace to space things instead of \&
%   and it did not use \& to align the following things on a line
%   and it did not use multiple aligned into the gather
%   (what's the point of gather if you don't have several aligned inside? use a single align* instead)
% }
% oh... that's good to know! the tech report is also riddled with this... :/

Above is a fraction of the shared syntax of all the programming languages of this paper.
\MP{why a fraction? what exprs are missing}
%% MK: Things that are missing here are: Evaluation contexts, states Ω and transitively other definitions affected by this
The syntax and semantics are mostly standard and therefore omitted.
However, when allocating arrays with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the cells and $\expr[_{2}]$ the overall size of an array.

\MPin{
	below is semantics info. should go with their symbols and the semantics rules.
	and a better description of the semantics symbols.
% MK: this moved to appendix! should we bring it back?
}
The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
Because of the sandboxing, locations pointing into heaps must be associated with a control-tag $\sandboxtag$ (either $\ctx$ or $\comp$) to know the heap they point into.
Other metadata in $\Omega$ entails a list of functions with information on whether they belong to the context or the component.
That way, when calling a function, the semantics can take care of emitting appropriate events that signal a context switch.
For the languages considered in this paper, functions are second-class.
So, the only values are pairs, pointers, and natural numbers.
Another non-standard part of the languages is the concept of the poison-tag.
The poison-tag marks an array as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
Because of this, it is possible that a program like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ yields a trace $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$ when executing.

@@

Possible events that are emitted during execution are similar to the ones of the specification trace model (\Cref{subsec:propdefs}).
The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.\MP{why? to help with proofs?}
Hereby, a $\ev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context- to component-level.
Contrary, $\ev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\ev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component- to context-level.
The environmental semantics emits the tag $\nocomm$ for calls from context- to context-level or component- to component-level.
In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.
\MP{intuition? these are aux functions for the semantics}
